---
title: "hw06 Data wrangling wrap up"
output: github_document
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#### Load data and packages

```{r}
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(MASS))
suppressPackageStartupMessages(library(singer))
suppressPackageStartupMessages(library(ggmap))
suppressPackageStartupMessages(library(repurrrsive))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(leaflet))

```


#### 2. Writing functions
Write one (or more) functions that do something useful to pieces of the Gapminder or Singer data.

For this task, I use the China's data from gapminder dataframe to explore it. So first we get the data for China.
```{r}
china <- "China" 
(china_dat <- gapminder %>% 
  filter(country == china))
```

Then we can use the linear model to fit the liefExp versus the year. 

Firstly, we create a linear model which returns the coefficient of the intercept and the slope. 

```{r}
linearModel <- function (dat, offset = 1952) {
  #fit a linear regression model
  linMod <- lm(lifeExp ~ I(year - offset), data = dat)
  returnDat <- c(linMod$coef)
  return(setNames(data.frame(t(returnDat)), c("intercept", "slope")))
}
```

We can directly apply this model the China data. 
```{r}
linearModel(china_dat)
```

```{r}
p <- ggplot(china_dat, aes(x = year, y = lifeExp))
p + geom_point() + geom_smooth(method = "lm", se = FALSE)
```

So we can see that the linear regression model doesn't works well for the China data.

Since the lifeExp doesn't change lineary with the year, we consider creating a quadratic model which may better fit the data. The model is created below. 

```{r}
quadraticModel <- function (dat, offset = 1952) {
  #fit a quadratic regression model
  quadraticMod <- lm(lifeExp ~ I(year - offset), data = dat)
  returnDat <- c(quadraticMod$coef)
  return(setNames(data.frame(t(returnDat)), c("intercept", "slope")))
}
```

Then, we can apply this model to the China data.
```{r}
quadraticModel(china_dat)
```

Actually, the quadratic model have a better result for this data. 
```{r}
p <- ggplot(china_dat, aes(x = year, y = lifeExp))
p + geom_point() + geom_smooth(method = "lm", formula = y ~ x + I(x^2), se = FALSE)
```

Apart from the quadratic regression model, we may also wish to fit a robust regression model. Similarly, we can write another function for the robust regression case.
```{r}
robustModel <- function (dat, offset = 1952){
  #fit a robust regression model
  robMod<- rlm(lifeExp ~ I(year - offset), data = dat)
  returnDat <- c(robMod$coef)
  return(setNames(data.frame(t(returnDat)), c("intercept", "slope")))
}
```


```{r}
robustModel(china_dat)
```

These result are very similar to the linear model.

```{r}
p <- ggplot(china_dat, aes(x = year, y = lifeExp))
p + geom_point() + geom_smooth(method = "rlm", se = FALSE)
```


#### 4. Work with the singer data
##### 4.1 Use purrr to map latitude and longitude into human readable information on the band’s origin places. 

In order to access the location, we need first register the Google API. 
```{r}
register_google("AIzaSyCofrdT-5MetLRbrW9Rt-l50gAeEr3eCNY")
```

Here is the `singer_locations` data frame. 
```{r}
str(singer_locations)
```

```{r}
head(singer_locations) %>% 
  knitr::kable()
```

Let's take a look at only the cities and their latitude and longitude. 
```{r}
cities_location <- singer_locations %>% 
  dplyr::select(city, latitude, longitude)

head(cities_location, 10) %>%    
  knitr::kable()
```

As we can from the table, there are a lot of `NA` table cells. We can instaed remove the `NA`s from the dataframe using the following codes:

```{r}
cities_location_without_NA <- cities_location %>% 
  filter(!is.na(city) & !is.na(latitude) & !is.na(longitude)) # remove NA
  
head(cities_location_without_NA, 10) %>% 
  knitr::kable()  
```

For simplicity, we only consider the first 30 entries. 
```{r}
cities_location_without_NA_30 <- cities_location_without_NA %>% 
  slice(1:30)

ggmap_loc <- mapply( FUN = function(lon, lat) { 
    revgeocode(c(lon, lat), output = "address") 
  }, 
  cities_location_without_NA_30$longitude, cities_location_without_NA_30$latitude
)
```

Then, we can take a look at the head of this dataframe. 
```{r}
head(ggmap_loc, 30)
```

And we can compare these with the data extracted from ggmap.
```{r}
knitr::kable(cbind(ggmap_loc, cities_location_without_NA_30$city),
             col.names = c("cities extracted from ggmap", "cities extracted from singer_location"),
             align = "c")
```


##### 4.2 Try to check wether the place in city corresponds to the information you retrieved.

For this part, we can use `grepl` function to check whether the city extracted from singer_location in the cities in ggmap.

```{r}
mapply(grepl, cities_location_without_NA_30$city, ggmap_loc, fixed = TRUE) %>% 
  knitr::kable()  
```

We can see from the result that there are a lof of mismatch. This is due to the format of address. We can solve this issue as shown in the following code. 

```{r}
split_word <- function(list) { # helper function to split string 
   r <- str_split(list, pattern = boundary("word"))
  return(r)
}

ggmap_location <- split_word(ggmap_loc)
cities_location <- split_word(cities_location_without_NA_30$city)

mapply(intersect, cities_location, ggmap_location) %>% 
  lapply(length) %>% 
  map(function(len) len > 0)
```

The result still exsits some mismatch due to the orginal mismatch problem in the dataframe of ggamp and singer_location. 

##### 4.3 If you still have time, you can go visual: give a look to the library leaflet and plot some information about the bands. 

We can use the following code to generate a visual figure. 
```{r}
cities_location_without_NA_30 %>%  
  leaflet()  %>%   
  addTiles() %>%  
  addCircles(lng = cities_location_without_NA_30$longitude,
             lat = cities_location_without_NA_30$latitude,
             popup = cities_location_without_NA_30$city
            )
```

